---
title: "initiate-data"
output: html_document
---

Introduction
-----

```{r init}
here::i_am('code/initiate-data.Rmd')
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE)
```

This script outlines the process to automate the updating of a duckdb database for new file released by NEON. This process uses the `neonstore` package to manage data file download and storage for NEON data releases. The database of up-to-date files are stored within the `data/database-files/` folder. The updating of data files will be automated approximately every 6 months through github actions. This will allow locally running scripts to convert the data into usable formats for size spectra analysis.

To establish the initial database `neonstore` was used to download stream macroinvertebrate and fish data products. We first set the local database directory (e.g., `r here::here('data/database-files/')`) and pass this directly to the `neon_download()` function.

```{r package setup}
library(neonstore)
# source the update_ function
# source(here::here('code/update-data-products.R'))
```

Using the `neon_download()` function we can automate the download into a pre-existing folder using established credentials. In the automated script, this is handled with the env variable `${{ github.workspace}}/data/database-files`. The data products for download are controlled with the dps object. Then they are added to the database with `neon_store()`

```{r update inverts}

streamsites=c("HOPB", "LEWI", "POSE", "CUPE",
              "GUIL", "KING", "MCDI", "LECO",
              "WALK", "MAYF", "ARIK", "BLUE",
              "PRIN", "BLDE", "COMO", "WLOU",
              "SYCA", "REDB", "MART", "MCRA",
              "BIGC", "TECR", "OKSR", "CARI")

db_path = here::here('data/database-files/')

dps = c("DP1.20120.001","DP1.20107.001","DP1.20190.001")

neon_download(product = dps,
              dir = db_path)

neon_store(dir = here::here('data/database-files/'),
           db = neon_db(dir = db_path, read_only = FALSE))
              
```

Once these initial files are downloaded and added to the database, these objects are uploaded to the bucket

```{r initialize the s3 bucket}

# create a temporary directory for the export parquet files
export_dir = file.path(tempdir(), 'parquet')

# export the local parquet file structure to the local database
neon_export_db(dir = export_dir,
               db = neon_db(dir = here::here('data/database-files/')))

# we can check the files in the temporary directory for uploading
list.files(export_dir,
           recursive = TRUE,
           full.names = TRUE)
# set the s3 bucket specs 
bucket = "neonsizedata-392202703749-us-east-2-an"
prefix = "neonstore/parquet"

s3 = arrow::s3_bucket(bucket = paste0(bucket,"/",prefix,"/"))

neon_sync_db(s3 = s3,
             dir = export_dir)

# 
# # get the local files in the parquet export
# 
# db_files <- list.files(
#   export_dir,
#   recursive = TRUE,
#   full.names = TRUE
# )
# 
# # upload all the files, preserving the paths
# for(f in db_files){
#   rel <- sub(
#     paste0("^", normalizePath(export_dir, winslash = "/", mustWork = TRUE), "/?"),
#     "",
#     normalizePath(f, winslash = "/", mustWork = TRUE)
#   )
# 
# 
# s3_key = file.path(prefix, rel)
# 
# put_object(
#   file = f,
#   object = s3_key,
#   bucket = bucket
# )
# }

```

We can then view the files available on the s3 bucket

```{r view bucket}

get_bucket(bucket = bucket,
           prefix = prefix)
```

And we can connect to the s3 bucket to perform actions locally.

```{r connect remote bucket}

remote = neon_remote_db(bucket = s3)

```


```{r read fish files}
neon_remote(
  product = "DP1.20120.001",
  db = remote
  ) %>% 
  dplyr::slice_sample(n = 1, by = table) %>%
  dplyr::collect() %>% 
  View()
```

